%Chapter 2 Stuff
We thank Ryan Chornock, Maria Drout, Wen-fai Fong, Ryan Foley, Daniel Kasen, Brian Metzger, Armin Rest, and Ken Shen for helpful discussions and providing model data during the course of this analysis. P.S.C. is grateful for support provided by the NSF through the Graduate Research Fellowship Program, grant DGE1144152.

%Chapter 3 Stuff
The Berger Time-Domain Group at Harvard is supported in part by the NSF through grants AST-1411763 and AST-1714498, and by NASA through grants NNX15AE50G and NNX16AC22G. P.S.C. is grateful for support provided by the NSF
through the Graduate Research Fellowship Program, grant DGE1144152. The UCSC group is supported in part by NSF grant AST--1518052, the Gordon \& Betty Moore Foundation, and from fellowships from the Alfred P.\ Sloan Foundation and the David and Lucile Packard Foundation to R.J.F.

This work has made use of data from the European Space Agency (ESA)
mission {\it Gaia} (\url{https://www.cosmos.esa.int/gaia}), processed by
the {\it Gaia} Data Processing and Analysis Consortium (DPAC,
\url{https://www.cosmos.esa.int/web/gaia/dpac/consortium}). Funding
for the DPAC has been provided by national institutions, in particular
the institutions participating in the {\it Gaia} Multilateral Agreement.
This publication makes use of data products from the Two Micron All Sky Survey,
which is a joint project of the University of Massachusetts and the Infrared Processing
and Analysis Center/California Institute of Technology, funded by the National Aeronautics
and Space Administration and the National Science Foundation. This publication makes use
of data products from the Wide-field Infrared Survey Explorer, which is a joint project of the
University of California, Los Angeles, and the Jet Propulsion Laboratory/California Institute of
Technology, funded by the National Aeronautics and Space Administration.

% Chapter 4 Stuff
P.S.C. is grateful for support provided by the NSF
through the Graduate Research Fellowship Program, grant DGE1144152.
R.J.F.\ gratefully acknowledges support from NSF grant AST--1518052 and
the Alfred P.\ Sloan Foundation. D.E.H.\ was supported by NSF CAREER grant
PHY-1151836. He also acknowledges support from the Kavli Institute for
Cosmological Physics at the University of Chicago through NSF grant
PHY-1125897 as well as an endowment from the Kavli Foundation.

This research uses services or data provided by the NOAO Science
Archive. NOAO is operated by the Association of Universities for
Research in Astronomy (AURA), Inc. under a cooperative agreement with
the National Science Foundation. The computations in this paper were
run on the Odyssey cluster supported by the FAS Division of Science,
Research Computing Group at Harvard University. This research has made
use of the NASA/IPAC Extragalactic Database (NED) which is operated by
the Jet Propulsion Laboratory, California Institute of Technology,
under contract with the National Aeronautics and Space Administration.
Light curve data for PS15cdi were obtained from The Open Supernova
Catalog \citep{Guillochon+17}. Some of the results in this paper have been
derived using the HEALPix package \citep{Gorski+05}.

Funding for the DES Projects has been provided by the DOE and NSF (USA),
MEC/MICINN/ MINECO (Spain), STFC (UK), HEFCE (UK). NCSA (UIUC),
KICP (U. Chicago), CCAPP (Ohio State), MIFPA (Texas A\&M), CNPQ, FAPERJ,
FINEP (Brazil), DFG (Germany) and the Collaborating Institutions in the Dark Energy Survey.

The Collaborating Institutions are Argonne Lab, UC Santa Cruz,
University of Cambridge, CIEMAT-Madrid, University of Chicago, University College London,
DES-Brazil Consortium, University of Edinburgh, ETH Z{\"u}rich, Fermilab,
University of Illinois, ICE (IEEC-CSIC), IFAE Barcelona, Lawrence Berkeley Lab,
LMU M{\"u}nchen and the associated Excellence Cluster Universe,
University of Michigan, NOAO, University of Nottingham, Ohio State University, University of
Pennsylvania, University of Portsmouth, SLAC National Lab, Stanford University,
University of Sussex, Texas A\&M University, and the OzDES Membership Consortium.

The DES Data Management System is supported by the NSF under
Grant Number AST-1138766. The DES participants from Spanish institutions are partially
supported by MINECO under grants AYA2012-39559, ESP2013-48274, FPA2013-47986,
and Centro de Excelencia Severo Ochoa SEV-2012-0234. Research leading
to these results has received funding from the ERC under the EU's 7$^{\rm th}$
Framework Programme including grants ERC 240672, 291329 and 306478.

% Chapter 5
The Berger Time-Domain Group at Harvard is supported in part by the NSF through grants AST-1411763 and AST-1714498, and by NASA through grants NNX15AE50G and NNX16AC22G. PSC is grateful for support provided by the NSF through the Graduate Research Fellowship Program, grant DGE 1144152. VAV acknowledges support by the National Science Foundation through a Graduate Research Fellowship. DAB is supposed by NSF award PHY-1707954. We thank the University of Copenhagen, DARK Cosmology Centre, and the Niels Bohr International Academy for hosting R.J.F.\ during the discovery of GW170817/SSS17a, where he was participating in the Kavli Summer Program in Astrophysics, ``Astrophysics with gravitational wave detections."  This program was supported by the Kavli Foundation, Danish National Research Foundation, the Niels Bohr International Academy, and the DARK Cosmology Centre. The UCSC group is supported in part by NSF grant AST--1518052, the Gordon \& Betty Moore Foundation, the Heising-Simons Foundation, generous donations from many individuals through a UCSC Giving Day grant, and from fellowships from the Alfred P.\ Sloan Foundation and the David and Lucile Packard Foundation to R.J.F. EB acknowledges financial support from the European Research Council (ERC-StG-335936, CLUSTERS). DK \& EQ were funded in part by the Gordon and Betty Moore Foundation through Grant GBMF5076.

We are grateful for the heroic efforts of the entire staff at Gemini-South to continue obtaining observations of this target in evening twilight at high airmass as the object was setting.

This project used data obtained with the Dark Energy Camera (DECam), which was constructed by the Dark Energy Survey (DES) collaboration. Funding for the DES Projects has been provided by the DOE and NSF (USA), MISE (Spain), STFC (UK), HEFCE (UK), NCSA (UIUC), KICP (U. Chicago), CCAPP (Ohio State), MIFPA (Texas A\&M), CNPQ, FAPERJ, FINEP (Brazil), MINECO (Spain), DFG (Germany) and the collaborating institutions in the Dark Energy Survey, which are Argonne Lab, UC Santa Cruz, University of Cambridge, CIEMAT-Madrid, University of Chicago, University College London, DES-Brazil Consortium, University of Edinburgh, ETH Z{\"u}rich, Fermilab, University of Illinois, ICE (IEEC-CSIC), IFAE Barcelona, Lawrence Berkeley Lab, LMU M{\"u}nchen and the associated Excellence Cluster Universe, University of Michigan, NOAO, University of Nottingham, Ohio State University, University of Pennsylvania, University of Portsmouth, SLAC National Lab, Stanford University, University of Sussex, and Texas A\&M University.

Based in part on observations at Cerro Tololo Inter-American Observatory, National Optical Astronomy Observatory (NOAO Prop. ID: 2017B-0110, PI: E. Berger), which is operated by the Association of Universities for Research in Astronomy (AURA) under a cooperative agreement with the National Science Foundation

Based in part on observations obtained at the Gemini Observatory (Program IDs GS-2017B-Q-8 and GS-2017B-DD-4; PI: Chornock), which is operated by the Association of Universities for Research in Astronomy, Inc., under a cooperative agreement with the NSF on behalf of the Gemini partnership: the National Science Foundation (United States), the National Research Council (Canada), CONICYT (Chile), Ministerio de Ciencia, Tecnolog\'{i}a e Innovaci\'{o}n Productiva (Argentina), and Minist\'{e}rio da Ci\^{e}ncia, Tecnologia e Inova\c{c}\~{a}o (Brazil).
